### 2019-03-13 Proposal.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage[overload]{empheq}
\usepackage{float}
\usepackage{graphicx}
\usepackage{indentfirst}

%==== templates ====
\iffalse
%inserting code
\lstinputlisting[style=Matlab-editor]{test.m}

%insertng image
\begin{figure}[H]
\centering
\includegraphics[width=0.5\textwidth]{plot}
\caption{Blah}
\label{blahblah}
\end{figure}

%interting table
\begin{table}[H]
\centering
\begin{tabular}{|c c c c|} 
 \hline
 Col1 & Col2 & Col2 & Col3 \\ 
 \hline
 1 & 6 & 87837 & 787 \\ 
 2 & 7 & 78 & 5415 \\
 3 & 545 & 778 & 7507 \\
 4 & 545 & 18744 & 7560 \\
 5 & 88 & 788 & 6344 \\
 \hline
 \end{tabular}
\end{table}
\fi

%==== document specific commands ====

\renewcommand{\vec}[1]{\mathbf{#1}}
\DeclareRobustCommand{\qed}{\blacksquare}

\title{PHYSICS 212 Final Paper Proposal}
\author{Jixun Ding (SUID: 3624401)}
\date{Feb 8, 2019}

\begin{document}

\maketitle

I would like to follow a mixture of  \#4 and \#5 of the suggested paper topics. I would like to
\begin{itemize}
\item Study the Monte Carlo Renormalization Group (MCRG) and understand how it works.

\item Write an implementation of MCRG for the Ising Model on a 2D square lattice, find the critical exponents, and compare my results to original work in literature\cite{Swendsen1979}.

\item  Write an implementation of MCRG for the $O(3)$ Heisenberg model on a 2D square lattice and reproduce various results in original work in literature\cite{Shenker1980}. Confirm the model does not have a phase transition.

\end{itemize}

So far I have found: earliest formulation of MCRG by Ma\cite{Ma1976}; reviews of the methodology and results of MCRG by Gupta\cite{Gupta1986,Gupta1987}; two descriptions of MCRG by Swendsen\cite{Swendsen1982,Swendsen1984}. Textbook by Binder\cite{Binder2014} provide general background on Monte Carlo methods applied to statistical physics problems and a modern perspective on MCRG.

\bibliographystyle{siam}
\bibliography{proposalref}

\end{document}